# Factor Practice {#factorpractice}

```{r include=FALSE}
library(tidyverse)
```

## The basic structure

Let's say that you have 100 cups of sizes small, medium, or large. 

```{r}
cups <- c("small", "medium", "large")
set.seed(2020)
manyCups <- sample(cups, size = 100, replace = TRUE)
table(manyCups)
```

Note that `table()` renders these in alphabetical order. It's usually more useful to render them in order of magnitude by setting levels. 

```{r}
sizesCups <- factor(manyCups, levels = c("small", "medium", "large"))
sizesCups
table(sizesCups)
```

The **forcats** package is designed to work with factors. Although I use it frequently, I haven't generalized the ideas, or internalized them. That's a mistake. The RStudio cheat sheet is a nice place to begin with this: <https://rstudio.com/resources/cheatsheets/>. 

Let's use my favorite **starwars** illustration to help illustrate some of **forcats**'s usefulness.

```{r}
starwars %>% 
  filter(!is.na(species)) %>% 
  count(species, sort = TRUE)
```

We see that there are a ton of unique species in this list. 

```{r}
fct_count(starwars$species, sort = TRUE)

```

