# By Any Other Name {#changenames}

> A rose by any other name would smell as sweet.
> `r tufte::quote_footer('--- Bill Shakespeare')`

The names we're talking about here are column names. As opposed to row names, which are better accomplished as a separate column of categorical data, column names are encouraged, and vital. ISLR refers to these as output and input variables (also known as predictors, independent variables, features). Data basing thinks in terms of fields. 

This deceptively simple-seeming idea of changing column names gets complex quickly. The following YouTube by Richard Webster offers a nice description of the process: <https://www.youtube.com/watch?v=Okc0IL5uTnA>

```{r}
my.data <- data.frame(colOne=1:3, column2=4:6, column_3=7:9)
rownames(my.data) <- c("ant", "bee", "cat")
names(my.data)
colnames(my.data)

#make some changes
names(my.data) <- c("col_1", "col_2", "col_3")
my.data
names(my.data)[3] <- "col.3"
my.data

names(my.data)[names(my.data)=="col_2"]
my.data["col_2"]
my.data$col_2
my.data[,2]

names(my.data)[names(my.data)=="col_2"] <- "col.2"
my.data

names(my.data) <- gsub("_", ".", names(my.data))
my.data

rownames(my.data)
my.data$species <- rownames(my.data)
my.data
rownames(my.data) <- NULL
my.data
colnames(my.data) <- c("good", "better", "best", "species")
my.data

keep <- 2:ncol(my.data)
my.data[,keep]
```

## A financial example

Dealing with financial data often means cleaning variable names and getting pesky dollar signs and commas out of downloaded files. Check out this example:
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

First, load the **tidyverse**


```{r}
etf_data <- tribble(
  ~Sign, ~"Market Value",
  "VTI", "$172.22",
  "VXUS", "$52.99"
)

etf_data

# First let's address the crappy variable names:

etf_data_namedOK <- janitor::clean_names(etf_data)

etf_data_namedOK

# Now remove the offending dollar signs and commas (if present)

etf_data_namedOK$market_value <- as.numeric(gsub('\\$|,','',as.character(etf_data_namedOK$market_value)))

etf_data_namedOK

# Don't be alarmed by the rounding behavior in the console. You can see that the data is intact. 
str(etf_data_namedOK)

glimpse(etf_data_namedOK)

etf_data_namedOK$market_value

```

