# Data Exploration {#dataexploration}

Data exploration is one of the most important aspects of data science and forms the cornerstone of my drills. Nonetheless, I have lots of room for improvement.

I like Hadley Wickham's writing and find his approach exceptionally clear. Therefore, I'll use the **tidyverse**. 

```{r, message=FALSE}
library(tidyverse)
```

## Counting things. The naming of parts. 

```{r starfig-1a, fig.cap='Starwars Figure 1'}
starwars %>% 
  filter(!is.na(species)) %>% 
  count(species = fct_lump(species, 5), sort = TRUE) %>% 
  mutate(species = fct_reorder(species, n)) %>% 
  ggplot(aes(species, n)) + 
  geom_col() + coord_flip()
```

I like stacked bars for their economy, but it's easy to over do it. Supperimposing gender onto the columns seems easy...

```{r starfig-2, fig.cap='Starwars Figure 2'}
starwars %>% 
  filter(!is.na(species)) %>% 
  count(species = fct_lump(species, 5), gender = fct_lump(gender, 2), sort = TRUE) %>% 
  mutate(species = fct_reorder(species, n)) %>% 
  ggplot(aes(species, n, fill = gender)) + 
  geom_col() + coord_flip()
```

But note that I've got a problem: the Droids, which outnumber the Gungans, are now reordered to _after_ the Gungans. This happens because the $n$ that we're counting comprises subcategories of species _and_ gender.  Only three Gungan males exist (and no females), but that is enough to tie the Droid NA category. The Droid NA category come after the Gungan category, presumably because _male_ comes before _NA_, or because NA comes last (more likely). 

Exploring this, I see that I'm getting warning messages about the implicit NA's in gender. Note that the following renders a slightly different plot. I _still_ have not fixed the order of the species. 


```{r starfig-3, fig.cap='Starwars Figure 3'}
starwars %>% 
  filter(!is.na(species)) %>% 
  count(species = fct_lump(species, 5), gender = fct_lump(gender, 2), sort = TRUE) %>% 
  mutate(gender = fct_explicit_na(gender),
         species = fct_reorder(species, n)) %>% 
  ggplot(aes(species, n, fill = gender)) + 
  geom_col() + coord_flip()
```

The trick here is to use `group_by()` and `ungroup()` wisely. 

```{r starfig-4, fig.cap='Starwars Figure 4'}
starwars %>% filter(!is.na(species)) %>% 
  mutate(species = fct_lump(species, 5)) %>% 
  group_by(species) %>% 
  mutate(typeCount = n()) %>% 
  ungroup() %>% 
  mutate(species = fct_reorder(species, typeCount)) %>% 
  ggplot()+
  geom_bar(aes(species, fill = gender))+
  coord_flip()
```

As opposed to using `count()`, which progressively narrows the information available to be used, by using `group_by()`/`mutate()`/`ungroup()` with `geom_bar()` we have all of the variables still available for plotting. 

## fct_infreq

As expected, Hadley Wickham and Garrett Grolemund solve this more simply in **R4DS**:

```{r starfig-5, fig.cap='Starwars Figure 5. Serial mutates are used.'}
starwars %>% filter(!is.na(species)) %>% 
  mutate(species = fct_lump(species, 5)) %>% 
  mutate(species = species %>% fct_infreq() %>% fct_rev()) %>% 
  ggplot()+
  geom_bar(aes(species, fill = gender))+
  coord_flip()
```

Note that a single `mutate()` suffices to both lump factors, organize by frequency and reverse the order, as follows.

```{r starfig-6, fig.cap='Starwars Figure 6. A single mutate is used.'}
starwars %>% filter(!is.na(species)) %>% 
  mutate(species = fct_lump(species, 5) %>% fct_infreq() %>% fct_rev()) %>% 
  ggplot()+
  geom_bar(aes(species, fill = gender))+
  coord_flip()
```

## Weight weight, don't tell me!

One feature of `count()` is difficult for me to remember, mostly because it is *not* a count--it is a sum. 

```{r}
df <- tibble(
  Symbol = c("a", "b", "c", "c", "d", "e"),
  Value = 1:6
)

df %>% count(Symbol, sort = TRUE)
df %>% count(Symbol, wt = Value, sort = TRUE)
```

The above example shows that `count()` performs in the usual way unless we use the `wt =` argument, which then sums the value specified and generates `n`, which now indicates a weight, not a count of a nominal variable. 

## Summarize is another very useful function:

```{r}
starwars %>% 
  filter(!(is.na(species))) %>% 
  group_by(species) %>% 
  summarize(n=n(), mean = mean(height, na.rm = TRUE)) %>% 
  arrange(desc(n))
```

## Graphical displays

At the risk of biting off far more than I can chew without starting a separate chapter on my favorite data displays, I'd like to include a few plots not represented elsewhere.

### Joyplots

Joyplots are a little hokey but can be visually satisfying. 

```{r, message=FALSE}
library(ggjoy)
```
The following display is from data taken from **The Cancer Genome Atlas** in 2017. It combines glioblastoma with lower grade infiltrating gliomas. Since I'm only interested in diagnosis and patient age, that's all that I've assembled in the `CSV`. 

```{r}
DiffuseGlioma <- read_csv("data/tcga9.11.17.csv")

DiffuseGlioma <- na.omit(DiffuseGlioma) #gets rid of the NA's

head(DiffuseGlioma)

ggplot(DiffuseGlioma,aes(x=AGE,y=HISTOLOGY)) +
  geom_joy(rel_min_height = 0.01) +  # removes tails
  scale_y_discrete(expand = c(0.01, 0)) + # removes cutoff top
  labs(x="age (y)",y="number") +
  theme_minimal()
```

## Relative versus absolute risk

Relative risk is often used in medical literature to emphasize the importance of a finding--it often makes a result sound more impressive. However, when one considers absolute risk, the effect is obviously much smaller. 

The following is adopted from <https://www.r-bloggers.com/lying-with-statistics-one-beer-a-day-will-kill-you/>

**personograph** allows for cute displays of individuals in big groups

```{r, message=FALSE}
library(personograph)
```

Start with 2000 people. Some of them will have problems without alcohol exposure, about 18, in fact. The blogger choses 2000 people to start with because $0.7*18=1$

Note that this **doesn't** stratify for any other health problems, age, socio-economic status, etc. 

```{r}
n <- 2000
probl_wo_alc <- 18 / n

data <- list(first = probl_wo_alc, second = 1-probl_wo_alc)
personograph(data,  colors = list(first = "black", second = "#efefef"),
             fig.title = "18 of 2000 people with health problems",
             draw.legend = FALSE, n.icons = n, dimensions = c(20, 100), 
             plot.width = 0.97)
```

Now we illustrate the affect of 500 mL of alcohol per day. According to the Lancet article, the relative risk of serious illness following consumption of about 25 mL ethanol (500 mL beer at 5% ABV) increases by about 7%. 

```{r}
probl_w_alc <- 1 / n
 
data_2 <- list(first = probl_wo_alc, second = probl_w_alc, third = 1-(probl_wo_alc+probl_w_alc))
personograph(data_2, colors = list(first = "black", second = "red", third = "#efefef"),
             fig.title = "About 1 additional case with half a litre of beer per day",
             draw.legend = FALSE, n.icons = n, dimensions = c(20, 100),
             plot.width = 0.97)
```

