# Functions

```{r}
library(tidyverse)
```

From R4DS and the Solution book for R4DS:

```{r}
x <- c(1:10, rep(NA,3)) 

cV <- function(x){
  sd(x, na.rm = TRUE)/mean(x, na.rm = TRUE)
}

xVar <- function(x) {
  n <- length(x[!is.na(x)])
  m <- mean(x, na.rm = TRUE)
  sq_err <- (x - m)^2
  sum(sq_err, na.rm = TRUE) / (n - 1)
}
```

From Richie Cotton's DataCamp introductory course on function writing:

```{r}
roll_die <- function(n_die){
  die_sides <- 1:6
  sample(die_sides, n_die, replace = TRUE)
}

roll_die(6)
```

Let's apply some of these simple ideas to D&D. Say I play a character that can cast the *sleep* spell. This scales to different levels.

At first level, one rolls 5 six-sided die. 

```{r}
sleep <- function(){
  sum(sample(1:8, size = 5, replace = TRUE))
}

sleep()
```

If cast at higher levels, the spell adds 2 die per level above first: 7 die at second level, 9 die at third level, etc. Let's address this by adding a level variable to our function.

```{r}
sleep_scaled <- function(level){
  if(level > 1){add_die = (level-1)*2
  } else{
    add_die = 0
  }
  sum(sample(1:8, size = 5 + add_die, replace = TRUE))
}
```

A couple of things can be added to avoid the following goofy mishaps (or at least render useful error codes:

```{r}
sleep_scaled(0)

# sleep_scaled("rex")
```

First we load Richie Cotton's package {assertive}.

```{r}
library(assertive)
```

Then we get to work on the function:

```{r}
sleep_scaled <- function(level = 1){
  # assert_is_numeric(level)
  if(any(is_non_positive(level))){
    return("x contains nonpositive values, so the spell makes no sense")
  }
  if(level > 1){add_die = (level-1)*2
  } else{
    add_die = 0
  }
  sum(sample(1:8, size = 5 + add_die, replace = TRUE))
}

sleep_scaled(0)
sleep_scaled()
sleep_scaled(2)
# sleep_scaled("joseph")
```

Now, it would be useful to see the distribution of these rolls, to predict whether the spell would be effective.

```{r}
hist(replicate(10000, sleep_scaled()), breaks = 70)
```

