# Clustering {#clustering}

The page by Rohan Joseph does a really nice job illustrative hierarchical clustering:

<https://medium.com/@rohanjoseph_91119/learn-with-an-example-hierarchical-clustering-873b5b50890c>

Joseph uses a one dimensional dataset to show the difference between types of hierarchical clustering. Joseph's hand-drawn calculations actually make the distinctionb between single and complete linkage very well.

```{r, message=FALSE}
library(tidyverse)
```

```{r}
x <- c(7,10,20,28,35) #Joseph's 1D dataset
```

The first step is to make a distance matrix between the elements of the dataset. In one dimension, this is accomplished by subtraction. 

```{r}
d <- dist(x)
d
```

Before calling the **hclust** function, it's helpful to know the methods. 

  * Single Linkage :  merge in each step the two clusters, whose two closest members have the smallest distance.
  * Complete Linkage : merge the members of the clusters in each step which provide the smallest maximum pairwise distance.

```{r}
hclust(d)
plot(hclust(d))
```

Note that the default method is complete. 

```{r}
hclust(d, method = "single")
plot(hclust(d, method = "single"))
```



