# Regular expressions  {#regex}

```{r include=FALSE}
library(tidyverse)
```

## Seeking patterns

I'm not very good at manipulating regular expressions, so a page to practice manipulating them seems like a good idea. 

The RStudio cheat sheet is a nice place to start: <https://rstudio.com/resources/cheatsheets/>

Or Hadley Wickham and Garrett Grolemund's R4DS chapter on strings: <https://r4ds.had.co.nz/strings.html>

```{r}
string <- c("Hiphopopotamus", "Rhymenoceros", "time for bottomless lyrics")

pattern <- "t.m"

grep(pattern, string)

grep(pattern, string, value = TRUE)

grepl(pattern, string)

stringr::str_detect(string, pattern)

regexpr(pattern, string)
# This finds the starting position and the length of the 1st match

gregexpr(pattern, string)
# This finds starting position and length of all matches

stringr::str_locate(string, pattern)
# This finds starting and end position of first match

stringr::str_locate_all(string, pattern)
# This finds starting and end position of all matches
```

It seems like dates and combined alphanumeric strings are frequently subjected to the gaze of grep and company. 

```{r}
string2 <- c("123.abc", "09/28/2020", "ABC..123", "...")

grep(pattern = "[a-z]+", 
     x = string2)

grepl(pattern = "[a-z]+", 
     x = string2)
  
```

From the Intermediate R DataCamp course:

```{r}
rand_w <- c("A_2_4b3c", "Gi2/3_5")

sub(pattern = "[a-zA-Z]",
     replacement = "X",
     x = rand_w)
# sub replaces only the 1st instance of the match per string

gsub(pattern = "[a-zA-Z]",
     replacement = "X",
     x = rand_w)
# gsub replaces all instances of the match per string
```

**stringr** markedly alters the base R process. To get a handle on it, R4DS comes to the rescue. 

For instance. Special characters have to be escaped. These include, `"`, `'`, `\`, `\n`, `\t`. For a complete list of these characters, see `?"'"`. 

```{r}
string1 <- "This is a string"
double_quote <- "\"" # or '"'


x <- c("\"", "\\")
x

writeLines(x)

```

