# Modeling {#modeling}

```{r include=FALSE}
library(tidyverse)
library(rpart)
```

## Modeling notes

Notes on/inspired by the Machine Learning with caret in R DataCamp course. 


Now for a regression example from the course:

```{r}
library(caret)

model <- train(
  price~., 
  diamonds,
  method = "lm",
  trControl = trainControl(
    method = "cv", 
    number = 10,
    verboseIter = TRUE
  )
)

# Print model to console
model
```

Note that **caret** handles the work of splitting test sets and calculating RMSE. 

Another example from the DataCamp course.

```{r}
library(MASS)

# Fit lm model using 5-fold CV: model
model <- train(
  medv~., 
  Boston,
  method = "lm",
  trControl = trainControl(
    method = "cv", 
    number = 5,
    verboseIter = TRUE
  )
)

# Print model to console
model
```

Cross validation can *itself* be repeated. The following is a 5-fold cross validation repeated 5 times.

```{r}
# Fit lm model using 5 x 5-fold CV: model
model <- train(
  medv ~ ., 
  Boston,
  method = "lm",
  trControl = trainControl(
    method = "repeatedcv", 
    number = 5,
    repeats = 5, 
    verboseIter = TRUE
  )
)

# Print model to console
model

predict(model, Boston)
```

## Classification

### An **rpart** classification example.

```{r}
model <- rpart(
  Species ~ .,
  data = iris,
  method = "class"
)

predicted <- predict(model, iris, type = "class")
predicted
mean(predicted == iris$Species)
```

Study what's going on in **predict()**:

```{r}
?predict.rpart
```

## From the DataCamp course

```{r}
library(mlbench)
data("Sonar")

# First randomize the dataset.
rows <- sample(nrow(Sonar))
Sonar <- Sonar[rows,]

# The split it into testing and training sets. 
split <- round(nrow(Sonar)*0.60)
train <- Sonar[1:split,]
test <- Sonar[(split+1):nrow(Sonar),]

nrow(train)/nrow(Sonar)
```

Let's build a model.

```{r}
model <- glm(
  Class ~ .,
  family = binomial(link = "logit"),
  train
)

p <- predict(model, test, type = "response")
summary(p)

p_class <- as.factor(ifelse(p > 0.1, "M", "R"))

table(p_class, test[["Class"]])

confusionMatrix(p_class, test[["Class"]])
```

