# Working with GitHub {#github}

## Basic idea

GitHub will no longer take a password, but needs a personal access token. 

## A couple invaluable websites
For creating a personal access token:
<https://docs.github.com/en/github/authenticating-to-github/keeping-your-account-and-data-secure/creating-a-personal-access-token>

For deploying a personal access token:
<https://happygitwithr.com/credential-caching.html#credential-caching>


I first explored my credentials with the following:

```{r}
library(gitcreds)

gitcreds_set()
```

This shows off my github password, with the right menu selection, but not the PAT. 

```{r}
library(credentials)

set_github_pat()
```


For Bill's Drills, I set up this PAT as directed. It's probably not a great idea to keep it in this document, but I'll play around with it until I understand the rules and make sure that my Github is still working. 
**************

Note that this resulted in the revocation of my PAT. Drat! This is readily solved by getting another PAT from GitHub. 

More research is necessary. See:

```{r}
?set_github_pat
```

On returning home, I encountered trouble with R. This seems to have been settled by the old copy-and-paste the error into my browser and figure out a workaround that involved a couple different methods. Now I've generated another PAT for my home computer, and am hoping that GitHub will be happy with me now.


