---
title: "Bill's Drills Book"
author: "William McDonald"
date: "Last edited `r Sys.Date()`"
output: bookdown::gitbook
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: https://github.com/wliammd/billsDrills
---
---
title: "Bill's Drills Book"
author: "William McDonald"
date: "Last edited `r Sys.Date()`"
output: bookdown::gitbook
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: https://github.com/wliammd/billsDrills
---

# Drills: Part of Every Healthy Intellectual Diet {#intro}

The goal of this book is to organize my R drills into reasonable chunks, the better to understand my strengths and weaknesses, and to plan new forays into data science.

<!--chapter:end:index.Rmd-->


# **bookdown** Tips for This Document {#bookdownplan}

Placeholder


## Basic conventions 
## Inserting pictures
## Referencing other parts of the document
## Referencing citations:
## Figures.   
## Working with Github

<!--chapter:end:02bookdown_tips.Rmd-->


# Shape of Data {#datashape}

Placeholder


## Making data frames
## Gather, spread, pivoting in the tidyverse
## Gathering steam...
## Spread your wings
## Missing Data: Ich vemisse Dich!
### You complete(me)
### fill(ing) in the gaps
## Pivoting to something new
## Larger structures

<!--chapter:end:03shapes.Rmd-->


# By Any Other Name {#changenames}

Placeholder


## A financial example

<!--chapter:end:04names.Rmd-->


# Factor Practice {#factorpractice}

Placeholder


## The basic structure
## Not all nominal data is a factor
## Making variables into factors
## Inspecting factors
## fct_lump()
## fct_infreq() and fct_rev()
## Additional practice
## Reordering Factors
### Improving legibility: change the linetype
### fct_reorder2(): another way to improve legibility

<!--chapter:end:05factors.Rmd-->


# Subsetting {#subset}

Placeholder


## Subsetting using brackets
## Subset using brackets by omitting the rows and columns we don’t want
## Subset using brackets in combination with the which() function and the %in% operator
## Subset using the subset() function
## Subset using dyplyr's filter() and select()

<!--chapter:end:06subsetting.Rmd-->


# Data Exploration {#dataexploration}

Placeholder


## Counting things. The naming of parts. 
## fct_infreq
## Weight weight, don't tell me!
## Summarize is another very useful function:
## Graphical displays
### Joyplots
## Relative versus absolute risk

<!--chapter:end:07data_explore.Rmd-->


# Sampling {#sampling}

Placeholder


## Think about throwing a bunch of dice.
## A keen way to divide up a dataset into testing and training components. 

<!--chapter:end:08sampling.Rmd-->


# Simulating data {#simulating}

Placeholder


## Sample() 
## replicate()
## sample() revisited
## generating fixed levels -------------------------------------------------
## generating numerical sequences
## seq_along() and seq_len(). 
## generating random data from a probability distribution 
## Normal distribution:
## Binomial distribution:
## Uniform distribution
## Sampling from multiple distributions (building in a "difference")
## The good stuff: building in a difference based on a categorical variable
## A demonstration of the Central Limit Theorem
## Overlaying normal curve on histogram
## Crossing trial

<!--chapter:end:09simulatingData.Rmd-->


# Functions

Placeholder


## A little Dungeons & Dragons (tm) to spice it up
## Geometric and harmonic means
## for loop
## case_when()
## Compare this with if_else()

<!--chapter:end:10functions.Rmd-->

# Correlation Plots {#correlation}

```{r, message=FALSE}
library(tidyverse)
```


```{r}
head(iris)
iris %>% select(-Species) %>% cor()

M <- iris %>% select(-Species) %>% cor(method = "kendall")
```

```{r}
corrplot::corrplot(M)
corrplot::corrplot(M, method = "color")
corrplot::corrplot(M, method = "color", type = "upper")
corrplot::corrplot(M, method = "color", type = "upper", order = "hclust")
corrplot::corrplot(M, method = "color", type = "upper", order = "hclust", addCoef.col = "black")
corrplot::corrplot(M, method = "color", type = "upper", order = "hclust", addCoef.col = "black", tl.col="black")
corrplot::corrplot(M, method = "color", type = "upper", order = "hclust", addCoef.col = "black", tl.col="black", tl.srt = 45)
```





<!--chapter:end:11correlation.Rmd-->


# Dimensionality Reduction {#dimensionalityreduction}

Placeholder


## Background
## Experiment in showing dimensionality reduction for pathologists
### Principle component analysis
### t-distributed stochastic neighbor embedding

<!--chapter:end:12dimension_reduction.Rmd-->


# Clustering {#clustering}

Placeholder



<!--chapter:end:13clustering.Rmd-->


# Regular expressions  {#regex}

Placeholder


## Seeking patterns

<!--chapter:end:14seeking_patterns.Rmd-->


# US_Census_Data_Kyle_Walker_Presentation_YouTube {#census}

Placeholder


## Variables
## Part 2: Wrangling Census Data with **tidyverse** Tools
### Normalizing the data with `mutate()`.
### `group_by()` and `summarize()` in census analysis
### Margin of error considerations
### Visualizing margins of error
### `get_estimates()` and how to use them
### `ggbeeswarm()` automates some jitter considerations

<!--chapter:end:15US_census_work.Rmd-->


# **K Nearest Neighbors Revisited: Cross Validation Added** Tips for This Document {#knncv}

Placeholder



<!--chapter:end:16knncv.Rmd-->


# Modeling {#modeling}

Placeholder


## Modeling notes
## Out of sample error example
## A regression example from the course:
## Classification
### An **rpart** classification example.
## From the DataCamp course
## Random forest with **caret**
## Random forest model

<!--chapter:end:17modeling.Rmd-->


# More modeling {#moremodeling}

Placeholder


## Split target from predictors
## Multiple preprocessing methods
## Max Kuhn on reusing a trainControl

<!--chapter:end:18modeling2.Rmd-->

# Working with GitHub {#github}

## Basic idea

GitHub will no longer take a password, but needs a personal access token. 

## A couple invaluable websites
For creating a personal access token:
<https://docs.github.com/en/github/authenticating-to-github/keeping-your-account-and-data-secure/creating-a-personal-access-token>

For deploying a personal access token:
<https://happygitwithr.com/credential-caching.html#credential-caching>


I first explored my credentials with the following:

```{r}
library(gitcreds)

gitcreds_set()
```

This shows off my github password, with the right menu selection, but not the PAT. 

```{r}
library(credentials)

set_github_pat()
```


For Bill's Drills, I set up this PAT as directed. It's probably not a great idea to keep it in this document, but I'll play around with it until I understand the rules and make sure that my Github is still working. 
**************

Note that this resulted in the revocation of my PAT. Drat! This is readily solved by getting another PAT from GitHub. 

More research is necessary. See:

```{r}
?set_github_pat
```



<!--chapter:end:19github.Rmd-->

`r if (knitr:::is_html_output()) '# References {-}'`

<!--chapter:end:99references.Rmd-->

